### 2017-06-02 R/oregon.Rmd
Updates to oregon.Rmd

```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime.csv",
                              GBDFile = "~/Downloads/gbd.csv")
```

Suppose we wanted to create a scenario in which the mean cycling time
and mean walking time are doubled.  We need only to change two means
in the model and nothing else.  To do so we could avoid creating input
files by simply updating the baseline object with a few new
parameters.  (note: this is the overall mean for the whole
population. From this, _ITHIM_ will create means for each age & gender
strata)

First we need to know the name of the parameters we want to update.
Currently the list of parameters resides in the help page for the
ITHIM class (`help("ITHIM")`).  Alternatively one can use the method
`getParameterNames`.

@@

Now we need to retrieve the walking and cycling means from the
baseline model.  To do so we use the method `getMeans`, which returns a vector of means.  To find a
list of available methods use `getMethodsITHIM` with no arguments.

@@

We can also pull out the mode-specific means from that vector and assign them to their own variables.

@@

## Active Travel Estimates for Portland (MSA 6442)

We use the National Household Transportation Survey (NHTS) data stored
as a _TravelSurvey_ R object to estimate mean walk and cycle times.

@@

>>>>

```{r misc320, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime.csv",
                              GBDFile = "~/Downloads/gbd.csv")
```

The contents of the _ITHIM_ object are displayed in summary when the
object is printed.  See below.

```{r show, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline
```

Currently the list of internal model parameters resides in the help page for the
ITHIM class (`help("ITHIM")`).  Alternatively one can use the method
`getParameterNames` to display the names.

@@

If we wanted to examine the overal means in the model we could use the
method `getMeans`, which returns a vector of means.  To find a list of
available methods use `getMethodsITHIM` with no arguments.

@@

## Creating a Scenario ITHIM object

Suppose we wanted to create a scenario in which the mean cycling time
and mean walking time are doubled.  We simply need to replace the
active transport file with one that has the desired values.

Sample Files
* [activeTransportTime_scenario.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/activeTransportTime_scenario.csv)

```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.scenario <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime_scenario.csv",
                              GBDFile = "~/Downloads/gbd.csv")
```

```{r misc7620, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
modelMeans <- rbind(cbind(rbind(getWalkTime(ITHIM.baseline), getCycleTime(ITHIM.baseline), getNonTravelMETs(ITHIM.baseline)),vision = "baseline"), cbind(rbind(getWalkTime(ITHIM.scenario), getCycleTime(ITHIM.scenario), getNonTravelMETs(ITHIM.scenario)),vision = "scenario"))

p <- ggplot(modelMeans, aes(x = ageClass, y = mean, fill = vision))
p + geom_bar(stat = "identity", position = "dodge") + facet_grid(type ~ sex, scales = "free_y")  + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

In the above example we needed only to change two means in the model
and nothing else.  To do so we could avoid creating input files by
simply updating the baseline object with a few new parameters.  (note:
this is the overall mean for the whole population. From this, _ITHIM_
will create means for each age & gender strata)

We can use the mode-specific means from the _meanVec_ vector and
assign them to their own variables.

@@

## Creating an ITHIM object for Portland (MSA 6442)

### Active Transport Means
We use the National Household Transportation Survey (NHTS) data stored
as a _TravelSurvey_ R object to estimate mean walk and cycle times.

@@

### Disease Burden

```{r gbd, eval = TRUE, echo = TRUE, fig.width = fig.width, fig.asp = 1/phi}
```
